Parallel Multi-objective Optimization on CPU Using Information Framework for Constructing Global Optimization Algorithms

### 2017-07-10 2017/URAL-PDC/u_pdc_paper.tex
u_pdc: tiny fixes

\begin{abstract}
In the present work, a parallel algorithm for the multi-objective optimization is considered. The
considered approach is based on the application of the information-statistical algorithm to some
reduced unicriterial problem, the set of the global optima in which coincides to the set of the
weakly efficient solutions of the initial multi-objective problem. The sequential version of this
method has been considered earlier. In the present work, the parallelization scheme based on
the characteristics, which is common for all information-statistical global search algorithms,
was applied to the sequential algorithm of the multi-objective optimization. Also, in the present
work one of the techniques of accounting for the local properties of the optimized function
allowing accelerating the convergence essentially has been considered for the multi-objective
method for the first time.

@@

To find the complete set of efficient or weakly efficient points is the most computationally
complex problem, however it provides a wide choice of possible compromised solutions. A
plenty of approaches to the finding of these sets has been developed: the application of the
genetic algorithms to the multi-objective problem directly~\cite{DebPratap2002}, the use of the
parametric convolutions reducing the problem to a series of the scalar global optimization
problems~\cite{GergelKozinov2017}, various approaches to the solving the problems with the
convex~ \cite{ZhangZuo2013} or linear~\cite{Benson1998} criteria. The main problem in the
using of majority of these methods is the absence of the theoretical guarantee of the uniform
convergence to the whole sought set of the Pareto-optimal solutions on a wide class of
problems including the ones with the non-convex multiextremal criteria. A method having this
property in the case, when all criteria satisfy the Lipschitz condition has been developed on the
basis of the information global search algorithm~\cite{markinStrongin1993}. In the present
paper, the parallel and accelerated by the local refinement technique~\cite{strOptBook}(chapter
3) modification of this method will be considered.

@@

As it has been shown in~\cite{markinStrongin1993}, the set of weakly efficient solutions of the
reduced problem (\(\ref{eq:problem}\)) coincides to the set of the globally optimal solutions of
the problem (\ref{eq:aux_problem}) i. e.
\begin{equation}
  \label{eq:s}
  S([0;1])=\{x\in [0;1]:\varphi(x)=\varphi^*\}
\end{equation}
Also, it has been shown in~\cite{markinStrongin1993} that the function \(\varphi(x)\) satisfies
the H\"older condition when the requirements (\ref{eq:holder}) are satisfied. Thus, the
information-statistical global search algorithm can be applied to the function \(\varphi(x)\) in
order to solve the problem (\ref{eq:aux_problem}). However, \(\varphi(x)\) is defined through
the operator \(\max\{...\}\), therefore, it is difficult to compute it directly. In
\cite{markinStrongin1993}, a modification of the classic information-statistical
algorithm~\cite{mixedAlg} is given, in which the values of \(\varphi(x)\) are computed
approximately. Let us consider the modified version of this algorithm. The modification
consists in the use of the local refinement technique described in~\cite{strOptBook}(chapter 3)
as well as in the parallelization based on the characteristics~\cite{strOptBook}(chapter 5).

@@

\textit{Step 6.} Perform the new trials at the points \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
\begin{equation}
  x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2} - \sign(z_{t_j} - z_{t_j-1})\frac{|z_{t_j} - z_{t_j-
1}|^n}{2r}
\end{equation}
All \(p\) trials within this step can be performed in parallel on \(p\) computer devices.

@@

      f_{2}\left(y\right) = \frac{(3y_1 - 2*y[1])}{2} + (y_1-y_2 + 1)\27 ^ 2{2} + 15\\
      f_{3}\left(y\right) =   + \sin(y_1^2 + y_2^2)\\

@@

Sandy
Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores

@@

\paragraph{Parallel method results.} In order to demonstrate the speedup in iterations, which
MOALR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
solved for the values \(p=1,2,4,8,16\). The parameres of the method were the following:
\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). For the problem (\ref{eq:schaffer2})
\(\varepsilon=0.001\). The number of iterations is presented in Table \ref{tab:exResultsIters}
(in the braces, the cardinalityof the set \(S\) is given) and the speedup in iterations is presented
in in Table \ref{tab:exResultsItersSpeedup}. As it is evident form Table
\ref{tab:exResultsIters}, the cardinality of the set \(S\) changed insuffiliently when varying \(p\)
i. e. the quality of the obtained estimates remained the same. At the same time, the number of
iterations decreased linearly with increasing value of \(p\) (Table
\ref{tab:exResultsItersSpeedup}). In fig. \ref{fig:s_estimation_all}, the examples of the
numerical solutions of the considered problems are given.

@@

\paragraph{Parallel method time speedup.} The objective functions of all problems listed in this
section are featured by low computational complexity. In order to demonstrate the possibility to
obtain a speedup in time for the problems with the hard-to-compute criteria, additinal intensive
froating point computations not affecting the resulting values were introduced into the criteria.
The speedups obtained are given in Table \ref{tab:exResultsTimeSpeedup}. In all cases, the
speedup in time was less than the speedip in iterations because the operation of the optimizaion
method itself takes a part of the execution time. However, in most cases the parallelizaion was
efficient enough even in the case of 16 CPU threads. For more computation costly criteria, one can
expect the increasing of the speedup in time for a large number of streams.

\begin{table}[ht]
  \centering
  \caption{Results of numerical experiments: speedup in time}
  \label{tab:exResultsTimeSpeedup}
  \begin{tabular}{|l|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|}
\hline
\textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
\cline{2-6}
&\(1\)(time, s) & \(2\) & \(4\) & \(8\) & \(16\)\\
\hline
Markin-Strongin & 104.47 & 1.97 & 3.65 & 6.79 & 9.90 \\
\hline
Fonseca and Fleming 2d & 118.95 & 1.85 & 2.81 & 5.79 & 6.40 \\
\hline
Fonseca and Fleming 3d & 554.45 & 1.51 & 4.14 & 8.05 & 10.69 \\
\hline
Viennet problem & - & 1.91 & 3.79 & 6.39 & 6.37\\
\hline
Poloni's problem & 336.74 & 1.82 & 3.64 & 6.98 & 10.60 \\
\hline
\end{tabular}
\end{table}

\section{Conclusion}
In the present work, a method for solving the multi-objective optimizaion problems allowing
finding a uniform estimate of the set of the weakly-optimal points has been considered. The
techniques of parallelization and of the acceleration of convergence common for all algorithms
of similar structure .have been applied. The numerical experiments performed have
demonstrated a speedup in convergence when using the local refinement. Also, the
parallelization based on the characteristics was found to be efficient for this method. The
speedup in iterations at the parallelization based on the characteristics appeared to be linear in
most cases as for the information methods of scalar optimization (see the results of the
parlallelization of the scalar optimization method, for example, in
\cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
expedient to apply sthe parallel MOALR in the problems of low dimensionality (1 to 3) with
the computation-costly criteria. Besides, the property of uniform convergence of the considered
algorithm to the whole set of the weakly-optimal solutions has been demonstrated on the
numerical  examples.
%
% ---- Bibliography ----
%
%\bibliographystyle{plain}
%\bibliographystyle{alpha}
\bibliographystyle{unsrt}
%\bibliographystyle{splncs03}
\bibliography{u_pdc_refs}

>>>>

\begin{abstract}
In the present work, a parallel algorithm for the multi-objective optimization is considered. The
considered approach is based on the application of the information-statistical algorithm to some
reduced scalar problem, the set of the global optima in which coincides to the set of the
weakly efficient solutions of the initial multi-objective problem. The sequential version of this
method has been considered earlier. In the present work, the parallelization scheme based on
the characteristics, which is common for all information global search algorithms,
was applied to the sequential algorithm of the multi-objective optimization. Also, in the present
work one of the techniques of accounting for the local properties of the optimized function
allowing accelerating the convergence essentially has been considered for the multi-objective
method for the first time.

@@

To find the complete set of efficient or weakly efficient points is the most computationally
complex problem, however it provides a wide choice of possible compromised solutions. A
plenty of approaches to the finding of these sets has been developed: the application of the
genetic algorithms to the multi-objective problem directly~\cite{DebPratap2002}, the use of the
parametric convolutions reducing the problem to a series of the scalar global optimization
problems~\cite{GergelKozinov2017}, various approaches to the solving the problems with the
convex~\cite{ZhangZuo2013} or linear~\cite{Benson1998} objectives. The main problem in the
using of majority of these methods is the absence of the theoretical guarantee of the uniform
convergence to the whole sought set of the Pareto-optimal solutions on a wide class of
problems including the ones with the non-convex multiextremal criteria. A method having this
property in the case, when all criteria satisfy the Lipschitz condition has been developed on the
basis of the information global search algorithm~\cite{markinStrongin1993}. In the present
paper, the parallel and accelerated by the local refinement technique~\cite{strOptBook}(chapter
3) modification of this method will be considered.

@@

As it has been shown in~\cite{markinStrongin1993}, the set of weakly efficient solutions of the
reduced problem (\ref{eq:problem}) coincides to the set of the globally optimal solutions of
the problem (\ref{eq:aux_problem}) i. e.
\begin{equation}
  \label{eq:s}
  S([0;1])=\{x\in [0;1]:\varphi(x)=\varphi^*\}
\end{equation}
Also, it has been shown in~\cite{markinStrongin1993} that the function \(\varphi(x)\) satisfies
the H\"older condition when the requirements (\ref{eq:holder}) are satisfied. Thus, the
information-statistical global search algorithm can be applied to the function \(\varphi(x)\) in
order to solve the problem (\ref{eq:aux_problem}). However, \(\varphi(x)\) is defined through
the operator \(\max\{...\}\), therefore, it is difficult to compute it directly. In
\cite{markinStrongin1993}, a modification of the classic information
global search algorithm~\cite{mixedAlg} is given, in which the values of \(\varphi(x)\) are computed
approximately. Let us consider the modified version of this algorithm. The modification
consists in the use of the local refinement technique described in~\cite{strOptBook}(chapter 3)
as well as in the parallelization based on the characteristics~\cite{strOptBook}(chapter 5).

@@

\textit{Step 6.} Perform the new trials at the points \(x^{k+j}\), \(1 \leqslant j \leqslant p\):
\begin{equation}
  x^{k+j}=\frac{x_{t_j}+x_{t_j-1}}{2} - \sign(z_{t_j} - z_{t_j-1})\frac{|z_{t_j} - z_{t_j-
1}|^n}{2r}
\end{equation}
All \(p\) trials within this step can be performed in parallel on \(p\) computing devices.

@@

      f_{2}\left(y\right) = \frac{(3y_1 - 2y_2 + 4)^2}{8} + \frac{(y_1-y_2 + 1)^2}{27} + 15\\
      f_{3}\left(y\right) = \frac{1}{y_1^2+y_2^2 + 1} -1.1 \exp\{-(y_1^2+y_2^2)\}\\

@@

Sandy Bridge E5-2660 2.2 GHz processors, 64 Gb RAM. The CPUs had 8 cores (i. e. total 16 cores

@@

\paragraph{Parallel method results.} In order to demonstrate the speedup in iterations, which
MOALR provides when \(p > 1\), all problems from Section \ref{sec:test_problems} have been
solved for the values \(p=1,2,4,8,16\). The parameres of the method were the following:
\(r=4.5\), \(\varepsilon=0.01\), \(\alpha=15\). The number of iterations is presented in Table \ref{tab:exResultsIters}
(in the braces, the cardinalityof the set \(S\) is given) and the speedup in iterations is presented
in in Table \ref{tab:exResultsItersSpeedup}. As it is evident form Table
\ref{tab:exResultsIters}, the cardinality of the set \(S\) changed insuffiliently when varying \(p\)
i. e. the quality of the obtained estimates remained the same. At the same time, the number of
iterations decreased linearly with increasing value of \(p\) (Table
\ref{tab:exResultsItersSpeedup}). In fig. \ref{fig:s_estimation_all}, the examples of the
numerical solutions of the considered problems are given.

@@

\paragraph{Parallel method time speedup.} The objective functions of all problems listed in this
section are featured by low computational complexity. In order to demonstrate the possibility to
obtain a speedup in time for the problems with the hard-to-compute criteria, additinal intensive
froating point computations not affecting the resulting values were introduced into the criteria.
The speedups obtained are given in Table \ref{tab:exResultsTimeSpeedup}. In all cases, the
speedup in time was less than the speedip in iterations because the operation of the optimizaion
method itself takes a part of the execution time. However, in most cases the parallelizaion was
efficient enough even in the case of 16 CPU threads. For more computation costly criteria, one can
expect the increasing of the speedup in time for a large number of threads.

\begin{table}[ht]
  \centering
  \caption{Results of numerical experiments: speedup in time}
  \label{tab:exResultsTimeSpeedup}
  \begin{tabular}{|l|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|K{1.5cm}|}
\hline
\textbf{Problem} & \multicolumn{5}{c|}{\(p\)}\\
\cline{2-6}
&\(1\)(time, s) & \(2\) & \(4\) & \(8\) & \(16\)\\
\hline
Markin-Strongin & 104.47 & 1.97 & 3.65 & 6.79 & 9.90 \\
\hline
Fonseca and Fleming 2d & 118.95 & 1.85 & 2.81 & 5.79 & 6.40 \\
\hline
Fonseca and Fleming 3d & 554.45 & 1.51 & 4.14 & 8.05 & 10.69 \\
\hline
Viennet problem & TBD & TBD & TBD & TBD & TBD\\
\hline
Poloni's problem & 336.74 & 1.82 & 3.64 & 6.98 & 10.60 \\
\hline
\end{tabular}
\end{table}

\section{Conclusion}
In the present work, a method for solving the multi-objective optimizaion problems allowing
finding a uniform estimate of the set of the weakly-optimal points has been considered. The
techniques of parallelization and of the acceleration of convergence common for all algorithms
of similar structure have been applied. The numerical experiments performed have
demonstrated a speedup in convergence when using the local refinement. Also, the
parallelization based on the characteristics was found to be efficient for this method. The
speedup in iterations at the parallelization based on the characteristics appeared to be linear in
most cases as for the information methods of scalar optimization (see the results of the
parlallelization of the scalar optimization method, for example, in
\cite{barkalovLebedef2016}). From the results of the experiments, one can conclude that it is
expedient to apply sthe parallel MOALR in the problems of low dimensionality (1 to 3) with
the computation-costly criteria. Besides, the property of uniform convergence of the considered
algorithm to the whole set of the weakly-optimal solutions has been demonstrated on the
numerical examples.
%
% ---- Bibliography ----
%
%\bibliographystyle{plain}
%\bibliographystyle{alpha}
\bibliographystyle{unsrt}
%\bibliographystyle{splncs03}
\bibliography{u_pdc_refs}
